QOCO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2023-10-14 paper-conference.tex
F_revision

\author{
		$\text{Iman Rahmati}^{\dagger}$, $\text{Hamed Shah-Mansouri}^{\ddagger}$, and $\text{Ali Movaghar}^{\dagger}$
		\\ 
		$^\dagger$Department of Computer Engineering, Sharif University of Technology, Tehran, Iran \\
		$^\ddagger$Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran \\
		email: {\{iman.rahmati, hamedsh, movaghar\}}@sharif.edu
	}

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), the focal point is the efficient computation task offloading problem, which plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand responsive and reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic load uncertainties. The decentralized determination of offloading decisions poses a challenging problem for each mobile device. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a markov decision process (MDP) to maximize the long-term user's QoE individually. We propose a QoE-oriented computation offloading (QOCO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other mobile devices. %We integrate advanced techniques, including Long Short-Term Memory (LSTM) and Dueling Double Deep Q-Network (DQN), to enhance the accuracy of QoE estimation in the algorithm. 
Through numerical studies, we evaluate mobile devices’ QoE as well as multiple key performance metrics. When evaluating the simulation results for 50 mobile devices and 5 edge nodes, it becomes evident that the QOCO algorithm efficiently exploits the computational capacities of edge nodes. Consequently, it can bring about a significant improvement in average QoE by at least 37\%. This improvement is achieved by increasing the number of completed tasks by 14\% and reducing task delay and energy consumption by 9\% and 6\%, respectively, when compared to an existing algorithm.
\end{abstract}

@@

\lettrine{M}{ \,EC} \cite{mao2017survey} has emerged as a promising technological solution to overcome the challenges faced by mobile devices (MDs) when performing high computational tasks, such as real-time data processing and artificial intelligence applications \cite{zhou2019edge} \cite{yousefpour2019all}. In spite of the technological advancements in MDs, their limitations in computing capacity and battery life, may lead to task drops, processing delays, and an overall poor user experience. By offloading intensive tasks to nearby edge nodes (ENs), MEC effectively empowers computation capability and reduces the delay and energy consumption. This improvement enhances the users' QoE, especially for time-sensitive computation tasks \cite{shah2018hierarchical}. Efficient task offloading in MEC is a complex optimization challenge due to the dynamic nature of the network and the variety of MDs and servers involved \cite{jiang2019toward}. In particular, determining the optimal offloading strategy, scheduling the tasks, and selecting the most suitable EN are main challenges that demand careful consideration.

Researchers have recently proposed several task offloading algorithms to tackle the aforementioned issues. Mao \textit{et al.} in \cite{mao2016dynamic} introduced a computation offloading algorithm for MEC systems.This scheme aims to reduce the MD's energy consumption while still meeting its computation delay constraints. In \cite{zhang2016energy}, Zhang \textit{et al.} proposed an offloading scheme for MEC in heterogeneous cellular networks, taking into account the energy-constrained MEC scenario and the heterogeneity of MDs when optimizing the offloading decisions. Bi \textit{et al.} in \cite{bi2018computation} proposed an algorithm to optimize decision-making about computation offloading and power transferring in a wireless-powered MEC. Ning \textit{et al.}, in \cite{ning2018cooperative}, introduced a heuristic algorithm designed to enhance the efficiency of the partial computation offloading model. This algorithm effectively adapts offloading decisions dynamically. The works in \cite{mao2016dynamic}-\cite{ning2018cooperative} primarily focus on quasi-static systems and seldom account for time-varying conditions and are not well-suited for dynamic systems. Furthermore, the uncertain requirements and sensitive latency properties of computation tasks in MEC systems with limited resources pose nontrivial challenges that can significantly impact computation offloading performance.

@@

To cope with the dynamic nature of the network, recent research has proposed several DRL-based algorithms.  DRL combines reinforcement learning principles with deep neural networks to determine optimal decision-making policies \cite{arulkumaran2017deep}. By capturing the dynamics of environments and learning strategies for accomplishing long-term objectives through repeated interactions, DRL can effectively tackle the challenges arising from the ever-changing nature of networks, MDs, and servers' heterogeneity. This ultimately improves the MD users' QoE. In \cite{huang2019deep}, Huang \textit{et al.} focus on a wireless-powered MEC network. They propose a DRL-based online approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. Liao \textit{et al.} in \cite{liao2023online} introduced a double reinforcement learning algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency to minimize both task computation delay and energy consumption. In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility. In this approach, each MD requires quality-of-service information from other MDs. Tang \textit{et al.}, in \cite{9253665}, investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. In \cite{chen2019energy}, Chen \textit{et al.} introduced an online algorithm with polynomial time complexity for offloading in MEC. This algorithm achieves a close-to-optimal energy consumption while ensuring that the delay is bounded. In \cite{dai2020edge}, Dai \textit{et al.} introduced the integration of action refinement into DRL and designed an algorithm to concurrently optimize resource allocation and computation offloading. Zhou \textit{et al.}, in \cite{zhou2021deep}, used an MDP to study MEC and model the interactions between the policy-system environment. They proposed a Q-learning approach to achieve optimal resource allocation strategies and computation offloading. In their work \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks.

Although DRL-based methods have demonstrated their effectiveness in handling network dynamics, task offloading still encounters several challenges that require further attention. QoE is a time-varying performance measure that reflects user satisfaction and is not affectedonly by delay as assumed in  \cite{zhao2019deep}-\cite{9253665}, but also by energy consumption. A more comprehensive approach is required to address the dynamic requirements of individual users in real-time scenarios with multiple MDs and ENs. Nevertheless, prior research in \cite{huang2019deep}-\cite{gao2022large} has failed to tackle some or all of these challenges.

@@

\begin{itemize}
	\item \textit{Task Offloading problem in the MEC System:} We formulate the task offloading problem as an MDP for time-sensitive tasks. This approach takes into account the dynamic nature of workloads at the ENs and concentrates on providing high performance in the MEC system while maximizing the long-term QoE.

	\item \textit{DRL-based offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption as QoE factors, to quantify the MDs' QoE. We propose QOCO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's battery level, our approach leverages DQN \cite{mnih2015human} and LSTM \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors.

	\item \textit{Performance Evaluation:} We conducted comprehensive experiments to evaluate QOCO's performance. The simulation results confirm that when compared with the potential game-based offloading algorithm (PGOA) in \cite{yang2018distributed} and several benchmark methods, our algorithm effectively utilizes the processing capabilities of MDs and ENs, resulting in a substantial improvement in the average QoE of the system.
\end{itemize}

@@

\section{System Model} 
\label{sec:latexhints}
% Required for proper example rendering in the compiled PDF
\newcount\LTGbeginlineexample
\newcount\LTGendlineexample
\newenvironment{ltgexample}%
{\LTGbeginlineexample=\numexpr\inputlineno+1\relax}%
{\LTGendlineexample=\numexpr\inputlineno-1\relax%
	%
	\tcbinputlisting{%
		listing only,
		listing file=\currfilepath,
		colback=green!5!white,
		colslot=green!25,
		coltitle=black!90,
		coltext=black!90,
		left=8mm,
		title=Corresponding \LaTeX{} code of \texttt{\currfilepath},
		listing options={%
			slot=none,
			language={[LaTeX]TeX},
			escapeinside={},
			firstline=\the\LTGbeginlineexample,
			lastline=\the\LTGendlineexample,
			firstnumber=\the\LTGbeginlineexample,
			basewidth=.5em,
			aboveskip=0mm,
			belowskip=0mm,
			numbers=left,
			xleftmargin=0mm,
			numberstyle=\tiny,
			numbersep=8pt%
		}
	}
}%
We investigate an MEC system consisting of a set of MDs denoted by $\mathcal{I} = \{1, 2, ..., I\}$, along with a set of ENs denoted by $\mathcal{J} = \{1, 2, ..., J\}$, where $I$ represents the number of MDs and $J$ represents the number of ENs. We regard time as a specific episode containing a series of time slots denoted by $\mathcal{T} = \{1, 2, \ldots, T\}$, with $T$ time slots, which represents a duration of $\tau$ seconds. As shown in Fig.~\ref{fig1}, we consider two separate queues for each MD to organize tasks for local processing or dispatching to ENs, operating in a first-in-first-out (FIFO) manner. The MD's scheduler is responsible for assigning newly arrived tasks to each of the queues at the beginning of the time slot. On the other hand, we assume that each EN $j \in \mathcal{J}$ consists of $i$ FIFO queues, where each queue corresponds to an MD $i \in \mathcal{I}$. When each task arrives at the ENs, it is enqueued in the corresponding MD's queue. %We assume that if a task is completed in a certain time slot, the next task in the queue will start its operation at the beginning of the next time slot.

We define $z_i(t)$ as the index assigned to the computation task arriving at MD $i \in \mathcal{I}$ in time slot $t \in \mathcal{T}$. Let $\lambda_i(t)$ denote the size of this task in bits. The size of task $z_i(t)$ is selected from a discrete set $\Lambda = \{\lambda_1, \lambda_2, \ldots, \lambda_{\theta}\}$, where $\theta$ represents the number of available values. Hence, $\lambda_i(t) \in \Lambda \cup \{0\}$. We also denote the task's processing density $z_i(t)$ as $\rho_i(t)$ indicates the number of CPU frequency required to complete the execution of a unit of the task. Furthermore, we denote the deadline of this task by $\Delta_i(t)$ which is the number of time slots that the task must be completed to not be dropped.
%At the beginning of time slot $t \in \mathcal{T}$, if MD $i \in \mathcal{I}$ has a newly arrived task, we define $z_i(t)$ as the unique index assigned to that task. Let $\lambda_i(t)$ be the number of newly arrived bits at the beginning of time slot $t \in \mathcal{T}$. If a new task $z_i(t)$ exists at the start of time slot $t$, then $\lambda_i(t)$ is equal to the size of task $z_i(t)$. Otherwise, $\lambda_i(t) = 0$. We consider $\rho_i(t)$ to represent the processing density of task $z_i(t)$, indicating the number of CPU cycles required to process a unit of the task.

We define two binary variables, $x_i(t)$ and $y_{i,j}(t),$ to determine the offloading decision and offloading target, respectively. Specifically, $x_i(t)$ indicates whether task $z_i(t)$ is assigned to the execution queue ($x_i(t) = 0$) or to the dispatch queue ($x_i(t) = 1$), and $y_{i,j}(t)$ indicates whether task $z_i(t)$ is offloaded to EN $j \in \mathcal{J}$. If the task is dispatched to EN $j$, we set $y_{i,j}(t) = 1$; otherwise, $y_{i,j}(t) = 0$.

@@

\subsection{State Space}
A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state at time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's battery level, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
\begin{alignat}{2}
\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
	\label{18}
\end{alignat}
where the vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1), j \in \mathcal{J})$ represents the queue length of MD $i$ in ENs at the previous time slot, which the MD can compute this vector according to \eqref{10}. Let $\phi_i(t)$ denote the battery level of MD $i$ at time slot $t$. Considering the power modes of Android MDs, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

@@

\subsection{Action Space}
The action space represents the agent's behavior and the decisions made by the agent. In this context, we define $\boldsymbol{a}_i(t)$ to denote the action taken by MD $i \in \mathcal{I}$ at time slot $t \in \mathcal{T}$. These actions involve two decision, (a) Offloading decision to determine whether to offload the task, and (b) Offloading target to determine the EN to send the offloaded tasks. Thus, the action of MD $i$ in time slot $t$ can be concisely expressed as the following action tuple:
\begin{alignat}{2}
	\boldsymbol{a}_i(t) = (x_i(t), \boldsymbol{y}_i(t)),
	\label{20}
\end{alignat}
where $\boldsymbol{y}_i(t) = (y_{i,j}(t), j \in \mathcal{J})$ is the selected EN for offloading this task. %Hence, the decision space of each MD is thus equal to $\mathcal{A} = \{0,1\}^{1+N}$, with $j$ representing the number of available ENs.

@@

$\mathcal{F}_i^t =\bigg \{ t' \bigg|\; t' \leq t,\; \lambda_i(t')>0, \; (1 - x_i(t')) \; l_i^{\text{C}}(t') \; + $ \vspace{-2mm}
\begin{equation}
	\hspace{20mm} x_i(t')\sum\limits_{\mathcal{J}} \sum\limits_{n= t'}^{t}\mathbbm{1}(z_{i,j}^{\text{E}}(n)=z_i(t'))  \; l_{i,j}^{\text{E}}(n) = t \bigg\}.
	\label{27}  
\end{equation}
Therefore, MD $i$ observes a set of QoEs $\{\boldsymbol{q}_i(t') \mid t' \in \mathcal{F}_i^t\}$ at the beginning of time slot $t+1$, where the set $\mathcal{F}_i^t$ for some $i \in \mathcal{I}$ can be empty. Subsequently, MD $i$ sends its experience $(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t), \boldsymbol{q}_i(t), \boldsymbol{s}_i(t+1))$ to EN $j_i$ for each task $z_i(t')$ in $t' \in \mathcal{F}_i^t$.

@@

\begin{table}
	\renewcommand{\arraystretch}{1.2}
	\captionsetup{name=TABLE}
	\caption{Simulation Parameters}
	\scalebox{1}{%
		\begin{tabular}{ l|l } 
			\toprule
			\textbf{Parameter} & \textbf{Value}  \\ 
			\midrule
			%Number of MDs $i$  & 50  \\ 
			%Number of EN $j$  & 5  \\ 
			%Number of episodes $\textit{Ep.}$  & 1000  \\ 
			%Number of time slots $T$  & 100 (with $\tau = 0.1$ second) \\ 
			Computation capacity of MD $f_i$  & 2.6 GHz \\ 
			Computation capacity of EN $f_j^{\text{E}}$ & 42.8 GHz  \\ 
			Transmission capacity of MD $r_{i,j}(t)$ & 14 Mbps  \\ 
			Task arrival rate & 250 Task/Sec  \\ 
			Size of task $\lambda_i(t)$  & $\{\text{1.0, 1.1, \ldots , 7.0}\}$ Mbits \\ 
			Required CPU cycles of task $\rho_i(t)$  & $\{\text{0.197,0.297,0.397}\}$G/Mbits \\ 
			Deadline of task $ \Delta_i$  & 10 time slots (1 Sec)\\ 
			Battery level of MD $\phi_i(t)$  & $\{\text{25, 50, 75}\}$ Percent \\ 
			Computation power of MD $p_i^{\text{C}}$ & \text{$10^{-27}(f_i)^3$}  \\ 
			Computation power of EN $p_j^{\text{E}}$ & 5 w  \\ 
			Transmission power of MD $p_i^{\text{T}}$ & 2.3 w  \\ 
			Standby power of MD $p_i^{\text{I}}$& 0.1 w  \\ 
	\end{tabular}}
	\label{table}
\end{table}

In these simulations, we first evaluate the QOCO algorithm for each of the influencing QoE factors, including the number of completed tasks, overall energy consumption, and average delay. Subsequently, we demonstrate the overall improvement of QOCO in terms of average QoE by comparing it against the following benchmark methods across multiple challenging scenarios.
\begin{enumerate}

>>>>

\author{
		Iman Rahmati, ~Hamed Shah-Mansouri, and Ali Movaghar
%		$^\dagger$Department of Computer Engineering, Sharif University of Technology, Tehran, Iran \\
%		$^\ddagger$Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran \\
%		email: {\{iman.rahmati, hamedsh, movaghar\}}@sharif.edu
	}

\thanks{I. Rahmati and A. Movaghar are with the Department of Computer Engineering, Sharif University of Technology, Tehran, Iran (e-mail:$\{\text{iman.rahmati, movaghar} \}$@sharif.edu). 
}

\thanks{H. Shah-Mansouri is with the Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran (email: hamedsh@sharif.edu)}

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand responsive and reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic load and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a decentralized QoE-oriented computation offloading (QOCO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. %We integrate advanced techniques, including Long Short-Term Memory (LSTM) and Dueling Double Deep Q-Network (DQN), to enhance the accuracy of QoE estimation in the algorithm. 
Through numerical studies, we evaluate the performance of QOCO. Simulation results validate that the QOCO algorithm efficiently exploits the computational resources of edge nodes. Consequently, it can complete 14\% more tasks and reduce task delay and energy consumption by 9\% and 6\%, respectively. These together contribute to a significant improvement of at least 37\% in average QoE compared to an existing algorithm.
\end{abstract}

@@

\lettrine{M}{ \,OBILE} edge computing (MEC)\cite{mao2017survey} has emerged as a promising technological solution to overcome the challenges faced by mobile devices (MDs) when performing high computational tasks, such as real-time data processing and artificial intelligence applications \cite{zhou2019edge} \cite{yousefpour2019all}. In spite of the MDs' technological advancements, their limited computing power and battery may lead to task drops, processing delays, and an overall poor user experience. By offloading intensive tasks to nearby edge nodes (ENs), MEC effectively empowers computation capability and reduces the delay and energy consumption. This improvement enhances the users' QoE, especially for time-sensitive computation tasks \cite{shah2018hierarchical}. Efficient task offloading in MEC is a complex optimization challenge due to the dynamic nature of the network and the variety of MDs and servers involved \cite{jiang2019toward}. In particular, determining the optimal offloading strategy, scheduling the tasks, and selecting the most suitable EN for task offloading are main the challenges that demand careful consideration.

Researchers have recently proposed several task offloading algorithms to tackle the aforementioned issues. Mao \textit{et al.} in \cite{mao2016dynamic} introduced a computation offloading algorithm for MEC systems. This scheme aims to reduce the MD's energy consumption while meeting the computation delay constraints. In \cite{zhang2016energy}, Zhang \textit{et al.} proposed an offloading scheme for MEC in heterogeneous cellular networks, taking into account the energy-constrained MEC scenario and the heterogeneity of MDs when optimizing the offloading decisions. Bi \textit{et al.} in \cite{bi2018computation} proposed an algorithm to optimize decision-making about computation offloading and power transferring in a wireless-powered MEC. Ning \textit{et al.} in \cite{ning2018cooperative} introduced a heuristic algorithm designed to enhance the efficiency of the partial computation offloading model. This algorithm effectively adapts offloading decisions dynamically. The works in \cite{mao2016dynamic}-\cite{ning2018cooperative} primarily focus on quasi-static systems and are not well-suited for dynamic systems and time-varying conditions. Furthermore, the uncertain requirements and sensitive latency properties of computation tasks in MEC systems with limited resources pose nontrivial challenges that can significantly impact computation offloading performance.

@@

To cope with the dynamic nature of the network, recent research has proposed several DRL-based algorithms.  DRL combines reinforcement learning principles with deep neural networks to determine optimal decision-making policies \cite{arulkumaran2017deep}. By capturing the dynamics of environments and learning strategies for accomplishing long-term objectives through repeated interactions, DRL can effectively tackle the challenges arising from the ever-changing nature of networks, MDs, and servers' heterogeneity. This ultimately improves the MD users' QoE. In \cite{huang2019deep}, Huang \textit{et al.} focused on a wireless-powered MEC network. They proposed a DRL-based online approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. Liao \textit{et al.} in \cite{liao2023online} introduced a double reinforcement learning algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency to minimize both task computation delay and energy consumption. In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility. In this approach, each MD requires quality-of-service information from other MDs. Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. In \cite{chen2019energy}, Chen \textit{et al.} introduced an online algorithm with polynomial time complexity for offloading in MEC. This algorithm achieves a close-to-optimal energy consumption while ensuring that the delay is bounded. In \cite{dai2020edge}, Dai \textit{et al.} introduced the integration of action refinement into DRL and designed an algorithm to concurrently optimize resource allocation and computation offloading. Zhou \textit{et al.} in \cite{zhou2021deep} used an MDP to study MEC and modeled the interactions between the policy-system environment. They proposed a Q-learning approach to achieve optimal resource allocation strategies and computation offloading. In\cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks.

Although DRL-based methods have demonstrated their effectiveness in handling network dynamics, task offloading still encounters several challenges that require further attention. QoE is a time-varying performance measure that reflects user satisfaction and is not affected only by delay as assumed in  \cite{zhao2019deep}-\cite{9253665}, but also by energy consumption. A more comprehensive approach is required to address the dynamic requirements of individual users in real-time scenarios with multiple MDs and ENs. Nevertheless, prior research in \cite{huang2019deep}-\cite{gao2022large} fail to tackle some or all of these challenges.

@@

\begin{itemize}
	\item \textit{Task Offloading Problem in the MEC System:} We formulate the task offloading problem as an MDP for time-sensitive tasks. This approach takes into account the dynamic nature of workloads at the ENs and concentrates on providing high performance in the MEC system while maximizing the long-term QoE.

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QOCO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's battery level, our approach leverages deep Q-network (DQN)\cite{mnih2015human} and long short-term memory(LSTM) \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors.

	\item \textit{Performance Evaluation:} We conduct comprehensive experiments to evaluate the QOCO's performance. The simulation results confirm that when compared with the potential game-based offloading algorithm (PGOA) in \cite{yang2018distributed} and several benchmark methods, our algorithm effectively utilizes the processing capabilities of MDs and ENs, resulting in a substantial improvement in the average QoE of the system.
\end{itemize}

@@

\section{System Model} 
\label{sec:latexhints}
% Required for proper example rendering in the compiled PDF
\newcount\LTGbeginlineexample
\newcount\LTGendlineexample
\newenvironment{ltgexample}%
{\LTGbeginlineexample=\numexpr\inputlineno+1\relax}%
{\LTGendlineexample=\numexpr\inputlineno-1\relax%
	%
	\tcbinputlisting{%
		listing only,
		listing file=\currfilepath,
		colback=green!5!white,
		colslot=green!25,
		coltitle=black!90,
		coltext=black!90,
		left=8mm,
		title=Corresponding \LaTeX{} code of \texttt{\currfilepath},
		listing options={%
			slot=none,
			language={[LaTeX]TeX},
			escapeinside={},
			firstline=\the\LTGbeginlineexample,
			lastline=\the\LTGendlineexample,
			firstnumber=\the\LTGbeginlineexample,
			basewidth=.5em,
			aboveskip=0mm,
			belowskip=0mm,
			numbers=left,
			xleftmargin=0mm,
			numberstyle=\tiny,
			numbersep=8pt%
		}
	}
}%
We investigate an MEC system consisting of a set of MDs denoted by $\mathcal{I} = \{1, 2, ..., I\}$, along with a set of ENs denoted by $\mathcal{J} = \{1, 2, ..., J\}$, where $I$ and $J$ represent the number of MDs and ENs, respectively. We regard time as a specific episode containing a series of time slots denoted by $\mathcal{T} = \{1, 2, \ldots, T\}$, with $T$ time slots,  each representing a duration of $\tau$ seconds. As shown in Fig.~\ref{fig1}, we consider two separate queues for each MD to organize tasks for local processing or dispatching to ENs, operating in a first-in-first-out (FIFO) manner. The MD's scheduler is responsible for assigning newly arrived tasks to each of the queues at the beginning of the time slot. On the other hand, we assume that each EN $j \in \mathcal{J}$ consists of $I$ FIFO queues, where each queue corresponds to an MD $i \in \mathcal{I}$. When each task arrives at an EN, it is enqueued in the corresponding MD's queue. %We assume that if a task is completed in a certain time slot, the next task in the queue will start its operation at the beginning of the next time slot.

We define $z_i(t)$ as the index assigned to the computation task arriving at MD $i \in \mathcal{I}$ in time slot $t \in \mathcal{T}$. Let $\lambda_i(t)$ denote the size of this task in bits. The size of task $z_i(t)$ is selected from a discrete set $\Lambda = \{\lambda_1, \lambda_2, \ldots, \lambda_{\theta}\}$, where $\theta$ represents the number of these values. Hence, $\lambda_i(t) \in \Lambda \cup \{0\}$ to consider the case that no task has arrived. We also denote the task's processing density as $\rho_i(t)$ that indicates the number of CPU cycles required to complete the execution of a unit of the task. Furthermore, we denote the deadline of this task by $\Delta_i(t)$ which is the number of time slots that the task must be completed to avoid being dropped.
%At the beginning of time slot $t \in \mathcal{T}$, if MD $i \in \mathcal{I}$ has a newly arrived task, we define $z_i(t)$ as the unique index assigned to that task. Let $\lambda_i(t)$ be the number of newly arrived bits at the beginning of time slot $t \in \mathcal{T}$. If a new task $z_i(t)$ exists at the start of time slot $t$, then $\lambda_i(t)$ is equal to the size of task $z_i(t)$. Otherwise, $\lambda_i(t) = 0$. We consider $\rho_i(t)$ to represent the processing density of task $z_i(t)$, indicating the number of CPU cycles required to process a unit of the task.

We define two binary variables, $x_i(t)$ and $y_{i,j}(t)$ for $i \in \mathcal{I}$ and $j \in \mathcal{J}$ to determine the offloading decision and offloading target, respectively. Specifically, $x_i(t)$ indicates whether task $z_i(t)$ is assigned to the execution queue ($x_i(t) = 0$) or to the dispatch queue ($x_i(t) = 1$), and $y_{i,j}(t)$ indicates whether task $z_i(t)$ is offloaded to EN $j \in \mathcal{J}$. If the task is dispatched to EN $j$, we set $y_{i,j}(t) = 1$; otherwise, $y_{i,j}(t) = 0$.

@@

\subsection{State Space}
A state in our MDP represents a conceptual space that comprehensively describes the state of an MD facing the environment. We represent the MD $i$'s state at time slot $t$ as vector $\boldsymbol{s}_i(t)$ that includes the newly arrived task size, the queues information, the MD's battery level, and the workload history at the ENs. The MD observers this vector at the beginning of each time slot. The vector $\boldsymbol{s}_i(t)$ is defined as follows:
\begin{alignat}{2}
\boldsymbol{s}_i(t) = \Big(\lambda_i(t), \delta_i^{\text{C}}(t), \delta_i^{\text{T}}(t), \boldsymbol{\eta}_i^{\text{E}}\hspace{-0.4mm}(\hspace{-0.4mm}t\hspace{-0.4mm}-\hspace{-0.4mm}1\hspace{-0.5mm}),\phi_i(t), \mathcal{H}(t) \Big),
	\label{18}
\end{alignat}
where the vector $\boldsymbol{\eta}_i^{\text{E}}(t-1) = (\eta_{i,j}^{\text{E}}(t-1))_{j \in \mathcal{J}}$ represents the queue length of MD $i$ in ENs at the previous time slot, which the MD can compute this vector according to \eqref{10}. Let $\phi_i(t)$ denote the battery level of MD $i$ at time slot $t$. Considering the power modes of Android MDs, $\phi_i(t)$ is derived from the discrete set $\Phi=\{\phi_1,\phi_2,\phi_3\}$, corresponding to ultra power-saving, power-saving, and performance modes, respectively.

@@

\subsection{Action Space}
The action space represents the agent's behavior and the decisions made by the agent. In this context, we define $\boldsymbol{a}_i(t)$ to denote the action taken by MD $i \in \mathcal{I}$ at time slot $t \in \mathcal{T}$. These actions involve two decision, (a) Offloading decision to determine whether to offload the task, and (b) Offloading target to determine the EN to send the offloaded tasks. Thus, the action of MD $i$ in time slot $t$ can be concisely expressed as the following action tuple:
\begin{alignat}{2}
	\boldsymbol{a}_i(t) = (x_i(t), \boldsymbol{y}_i(t)),
	\label{20}
\end{alignat}
where the vector $\boldsymbol{y}_i(t)=(y_{i,j}(t))_{j \in \mathcal{J}}$ represents the selected EN for offloading this task. %Hence, the decision space of each MD is thus equal to $\mathcal{A} = \{0,1\}^{1+N}$, with $j$ representing the number of available ENs.

@@

$\mathcal{F}_i^t =\bigg \{ t' \bigg|\; t' \leq t,\; \lambda_i(t')>0, \; (1 - x_i(t')) \; l_i^{\text{C}}(t') \; + $ \vspace{-2mm}
\begin{equation}
	\hspace{20mm} x_i(t')\sum\limits_{\mathcal{J}} \sum\limits_{n= t'}^{t}\mathbbm{1}(z_{i,j}^{\text{E}}(n)=z_i(t'))  \; l_{i,j}^{\text{E}}(n) = t \bigg\}.
	\label{27}  
	\nonumber
\end{equation}
Therefore, MD $i$ observes a set of QoEs $\{\boldsymbol{q}_i(t') \mid t' \in \mathcal{F}_i^t\}$ at the beginning of time slot $t+1$, where the set $\mathcal{F}_i^t$ for some $i \in \mathcal{I}$ can be empty. Subsequently, MD $i$ sends its experience $(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t), \boldsymbol{q}_i(t), \boldsymbol{s}_i(t+1))$ to EN $j_i$ for each task $z_i(t')$ in $t' \in \mathcal{F}_i^t$.

@@

\begin{table}
	\renewcommand{\arraystretch}{1.2}
	\captionsetup{name=TABLE}
	\caption{Simulation Parameters}
	\scalebox{1}{%
		\begin{tabular}{ l|l } 
			\toprule
			\textbf{Parameter} & \textbf{Value}  \\ 
			\midrule
			%Number of MDs $i$  & 50  \\ 
			%Number of EN $j$  & 5  \\ 
			%Number of episodes $\textit{Ep.}$  & 1000  \\ 
			%Number of time slots $T$  & 100 (with $\tau = 0.1$ second) \\ 
			Computation capacity of MD $f_i$  & 2.6 GHz \\ 
			Computation capacity of EN $f_j^{\text{E}}$ & 42.8 GHz  \\ 
			Transmission capacity of MD $r_{i,j}(t)$ & 14 Mbps  \\ 
			Task arrival rate & 250 Task/Sec  \\ 
			Size of task $\lambda_i(t)$  & $\{\text{1.0, 1.1, \ldots , 7.0}\}$ Mbits \\ 
			Required CPU cycles of task $\rho_i(t)$  & $\{\text{0.197,0.297,0.397}\}$G/Mbits \\ 
			Deadline of task $ \Delta_i$  & 10 time slots (1 Sec)\\ 
			Battery level of MD $\phi_i(t)$  & $\{\text{25, 50, 75}\}$ Percent \\ 
			Computation power of MD $p_i^{\text{C}}$ & \text{$10^{-27}(f_i)^3$}  \\ 
			Computation power of EN $p_j^{\text{E}}$ & 5 w  \\ 
			Transmission power of MD $p_i^{\text{T}}$ & 2.3 w  \\ 
			Standby power of MD $p_i^{\text{I}}$& 0.1 w  \vspace{1mm}\\ 
			\toprule
	\end{tabular}}
	\label{table}
\end{table}
In these simulations, we first evaluate the QOCO algorithm for each of the influencing QoE factors, including the number of completed tasks, overall energy consumption, and average delay. Subsequently, we demonstrate the overall improvement of QOCO in terms of average QoE by comparing it against the following benchmark methods across multiple challenging scenarios.
\begin{enumerate}
